Biases in Planet Occurrence Caused by Unresolved Binaries in Transit Surveys

### 2018-04-18 paper/review/response_1.txt

Referee comment:

Section 2, end of paragraph 2, For each of the 4 points regarding missed
transit signals, how often do they occur? (could add references).

  Response:

  We assume the referee is referring to paragraph 3.  All of these
  effects are quite common, both in the Kepler survey and ground-based
  surveys, though we are not aware of any source that would allow us
  to estimate the probabilities.  Since these "pitfalls" are not the
  topic of this paper, we do not want to attempt such a calculation
  here.

@@

Referee comment:

Section 3.2, following equations 4, you may want to change "stars" to "host
star".

  Response:

  Equation 4 is in Section 3.1 (not Section 3.2) and is not followed
  by a sentence with the word "stars". Perhaps the referee is
  referring to equation 14?  We do not understand why it would be an
  improvement here to replace "stars" with "host star", so we have not
  taken any action.

@@

Referee comment:

Section 4.1, when exploring the Howard et al. planet radius distribution, how
sensitive are the results to the choice of gamma? Do mathematical artifacts
here make the results insensitive to the choice of the value of gamma?

  Response:

  We addressed this point by considering the cases of γ=-2.5, and
  γ=-3.5, which bracket the result from Howard et al. and are well
  outside the quoted uncertainty range.  Applying Equations 28 and 30,
  we find:

@@

Referee comment:

Section 4.4, end of paragraph 1, Add Lopez and Fortney citation.

  Response:

  We have corrected this omission, and also added a citation to Owen &
  Wu (2013).

@@

Referee comment:

Section 4.4, paragraph 2, "radius distribution is assumed to have a
complete absence of planets with sizes between 1.5 and 2R_Earth",
might clarify that the "true" radius distribution is assumed to have
this feature after fitting the apparent distribution.

  Response:

  We have tried to clarify this passage, to distinguish between the
  apparent rate distribution measured by Fulton et al. (2017) and the
  distribution assumed in our numerical experiment.

@@

Referee comment:

Section 4.4, paragraph 2, "may be even more devoid of planets than it appears",
might want to emphasize again here that this conclusion is very dependent on
the true distribution model plugged into the formalism.

  Response:

  We have added a few sentences to make our point more clearly:
  "The quantitative results of this numerical experiment obviously
  depend on the input assumptions.  The more general point is that
  whenever there is a gap-like feature in the true rate density
  distribution, the effects of binarity will tend to fill it in.  If a
  gap is observed in an apparent distribution, the true distribution
  must have a gap at least as deep, if not deeper."

@@

Referee comment:

Section 5, Hot Jupiter paragraph, states that result of this study find that
unresolved binaries likely result in an overestimated of hot-Jupiter occurrence
rates, implied by the Fulton gap model in Section 4.4. Previously, in section
4.2, paragraph 3, it states the occurrence rate density of planets greater than
8 Earth radii are underestimated. The abstract just says net systematic error
of order 5%. It would help to include a table describing the different tests in
this study, and how parameters are varied (how knobs are turned).

  Response:

  We see now that in Section 5 (Hot Jupiter paragraph) we erroneously
  referred to Section 4.4; this should have been a reference to
  Section 4.5 (the gaussian radius distribution model).  We have
  corrected the error.

  We have also followed the referee's advice and added a table
  summarizing the case studies, as well as the parameters that are
  varied.

@@

Referee comment:

- Recommend using the term "uncertainty" instead of "errors" .

  Response:

  We appreciate the suggestion but we prefer to leave the text as it
  is.  The term "systematic error" is in more common use than
  "systematic uncertainty", by about a factor of 5 according to Google
  Ngram viewer.

@@

********************************************************************************

Referee comment:

- May want to comment on parameterizing the occurrence rate in terms of planet
  radius and insolation flux. The semi major axis will depend on which star in
  the binary the planet orbits.

  Response: We have added a "Caveats and Outlook" paragraph that
  touches on all three of these issues.

@@

- May want to comment on paper by Danley Hsu: Occurrence rates from
  Kepler using approximate Bayesian computation. You could speculate
  on what their data generating process would yield when plugged into
  this formalism.

  Response: The work by Hsu et al. does look interesting, but since we
  lack experience with their technique we do not wish to speculate on
  how to modify it to accommodate binarity.

@@

- May want to comment on the parameter space and telescopes for resolving
  binaries of various separations, etc. and the probability of being able to
  resolve binaries with followup.

  Response: Given that there are so many possible approaches and
  instruments for imaging, RV, and astrometry, a quantitative analysis
  of this issue would be beyond the scope of this article.  In the
  revision, we have simply acknowledged that the combination of Gaia
  and TESS data will lessen the severity of the problem, compared to
  the situation with Kepler.

Other:

* We have changed the ylabel in Figures 4b and 6b from "number of planets per
  star" to "number of planets per star per bin".

* We now mention Figure 2 in the text (cf. appended List of Changes)

### 2018-04-18 paper/main.tex
update manuscript w/ JNW comments

\title{ Biases in Planet Occurrence Caused by Unresolved Binaries in
        Transit Surveys }
%
\correspondingauthor{L. Bouma}
\email{luke@astro.princeton.edu}
%
\author[0000-0002-0514-5538]{L. G. Bouma}
\affiliation{
    Department of Astrophysical Sciences,
    Princeton University,
    4 Ivy Lane, Princeton, NJ 08540, USA}
\author[0000-0003-1298-9699]{K. Masuda}
\affiliation{
    Department of Astrophysical Sciences,
    Princeton University,
    4 Ivy Lane, Princeton, NJ 08540, USA}
\affiliation{
    NASA Sagan Fellow}
\author[0000-0002-4265-047X]{J. N. Winn}
\affiliation{
    Department of Astrophysical Sciences,
    Princeton University,
    4 Ivy Lane, Princeton, NJ 08540, USA}
%
%
\begin{abstract}
%
Wide-field surveys for transiting planets, such as the NASA {\it
Kepler} and {\it TESS} missions, are usually conducted without
knowing which stars have binary companions. Unresolved and
unrecognized binaries give rise to systematic 
\deleted{errors}\added{uncertainty} in planet
occurrence rates, including misclassified planets and mistakes in
completeness corrections.  The individual errors can have different
signs, making it difficult to anticipate the net effect on inferred
occurrence rates. Here we use simplified models of signal-to-noise
limited transit surveys to try and clarify the situation. We derive a
formula for the apparent occurrence rate density measured by an
observer who falsely assumes all stars are single. The formula depends
on the binary fraction; the mass function of the secondary stars; and
the true occurrence of planets around primaries, secondaries, and
single stars. It also takes into account the Malmquist bias by which
binaries are over-represented in flux-limited samples. Application of
the formula to an idealized {\it Kepler}-like survey shows that for
planets larger than 2\,$R_\oplus$, the net systematic error is of
order 5\%.  In particular, unrecognized binaries are unlikely to be
the reason for the apparent discrepancies between hot Jupiter
occurrence rates measured in different surveys.  For smaller planets
the errors are potentially larger: the occurrence of Earth-sized
planets could be overestimated by as much as 50\%. We also show that
whenever high-resolution imaging reveals a transit host star to be a
binary, the planet is usually more likely to orbit the primary star
than the secondary star.
%
\end{abstract}
%
\keywords{
    methods: data analysis ---
    planets and satellites: detection ---
    surveys}
%
%

@@

A lingering concern about these studies is that in most cases,
investigators have assumed that all of the targets in the survey
are single stars~\citep[\textit{e.g.},][]{
  howard_planet_2012,fressin_false_2013,
  dressing_occurrence_2015,burke_terrestrial_2015}. 
In reality, many of the sources that are monitored in a transit survey are
unresolved multiple-star systems, mainly binaries.  Unrecognized
binaries cause numerous systematic \deleted{errors}\added{uncertainties} in 
the planetary occurrence
rates.  For example, when there is a transiting planet around a star
in a binary, the additive constant light from the second star reduces
the fractional loss of light due to the planet.  This makes transit
signals harder to detect and lowers the number of detections.  On the
other hand, a binary system presents two opportunities to detect
transiting planets, which could increase the overall number of
detections.

@@

There are many potential pitfalls in this calculation.  Some genuine
transit signals are missed even if they formally exceed the
signal-to-noise threshold, because of the probabilistic nature of
transit detection.  Planets can be misclassified due to statistical
and systematic \deleted{errors}\added{uncertainties} in the catalogued 
properties of the stars.  Some
transit-like signals are spurious, arising from noise fluctuations or
failures of ``detrending'' the astrophysical or instrumental
variations in the photometric signal.  Poor angular resolution leads
to blends between eclipsing binary stars and other stars along nearly
the same line of sight, producing signals that mimic those of
transiting planets.

Here, though, we will focus exclusively on problems that arise from
the fact that many stars exist in unresolved binary systems.  We will
also focus on the errors in planet occurrence as a function of radius,
rather than orbital period. This is because when more than one transit
is detected (as is usually required by the surveyors), the orbital
periods can be measured without ambiguity regardless of whether the
host star is single or one member of a binary.  Even with this narrow
focus, there are numerous sources of error.  All three of the
quantities in Equation~\ref{eq:occ_rate_simple} are biased:
\begin{enumerate}
%    
    \item The number of detected planets, $N_{\det}$, is actually the
      number of detected planets that {\it appear} to have size $r$,
      orbiting stars that {\it appear} to have mass $M$ \added{and
      radius $R$}.  Whenever the planet-hosting star is part of a
      binary,
%    
    \begin{itemize}
        \item the planet's size could be misclassified because of the
          reduction in the amplitude of the photometric signal;
%        
        \item the host star's properties could be misclassified
          because its light is combined with a second star of a
          different spectral type.
%        
    \end{itemize}
%    
    \item The number of searchable stars, $N_\star$, is biased
%    
    \begin{itemize}
%        
        \item toward lower values, because it does not include all of
          the secondary stars that were inadvertently searched for
          transiting planets;
%        
        \item toward higher values, because some of the stars that
          appeared to be searchable are in fact binaries for which the
          amplitude of the photometric signal would have been reduced
          to an undetectable level.
%        
    \end{itemize}
%    
    \item The transit probability $p_{\rm tra}$ is biased because the
      planet-hosting star could be misclassified.  At fixed orbital
      period, the transit probability is proportional to
      $\rho^{-1/3}$, where $\rho$ is the stellar mean
      density~\citep{Winn2010}.  Therefore, any 
      \deleted{errors}\added{uncertainties} in determining
      the host star's mean density lead to 
      \deleted{errors}\added{uncertainties} in the correction
      for the transit probability.
%    
\end{enumerate}

@@

We can now gauge the size of the systematic 
\deleted{errors}\added{uncertainties} associated
with unresolved binaries, for \deleted{cases of interest.}\added{the cases of 
interest listed in Table~\ref{tab:case_study_toc}.}
Throughout this
section we assume $R\propto M$, and $L \propto M^{\alpha}$ where
$\alpha = 3.5$, as is roughly the case for main-sequence stars.
With these assumptions, Equations~\ref{eq:dil_1} and~\ref{eq:dil_2}
become
\begin{equation}
  \mathcal{D}_1
  =  (1+q^\alpha)^{1/2}~~{\rm and}~~\mathcal{D}_2
  = q(1+q^{-\alpha})^{1/2}.
\end{equation}

@@

Again assuming $\Gamma_0=\Gamma_1=\Gamma_2$, the apparent occurrence
rate density is 
\begin{align}
    %see annoying_q_factor_dec22.pdf for derivation
    \notag
    %\frac{\left.\Gamma\a(r\a)\right|_{r\a= r}}{\Gamma_0(r)} 
    \frac{\Gamma\a(r\a)}{\Gamma_0(r\a)} 
    &=
    \frac{1}{1+\mu}
    \left[
      1 + \frac{1}{\mathcal{N}_q} \frac{n\b}{n\s}
    \left(
      \int_0^1 {\rm d}q\,q^\beta (1+q^\alpha)^{\frac{\gamma+4}{2}} +
      \right.
      \right. \\
      %isn't latex great?
      &\quad\quad\quad\quad\quad\quad\quad\quad\quad\quad\quad
      \left.\left.
      \int_0^1 {\rm d}q\,q^{\beta+\gamma+\frac{5}{3}} 
      (1+q^\alpha)^{\frac{3}{2}}
      (1+q^{-\alpha})^{\frac{\gamma+1}{2}}
    \right)
    \right],
    \label{eq:powerlaw_vary_binary}
\end{align}
Adopting the realistic numerical values $n\b/n\s=0.79$, $\alpha =
3.5$, $\beta=0$, and $\gamma=-2.92$, the summed integrals in
Equation~\ref{eq:powerlaw_vary_binary} give $(\ldots)\approx 1.503$,
and $\Gamma\a/\Gamma_0 = 1.048$.  The apparent occurrence rate density
is 4.8\% higher than the true value.
\added{Exploring different values for the power-law index,
we find that $\Gamma\a/\Gamma_0$ lies between $0.99$ and 
$1.10$, for $\gamma=-3.5$ and $\gamma=-2.5$, respectively.}
We conclude that for planets around
Sun-like stars with orbits within 0.25\,AU and radii in the range
2--17\,$R_\oplus$, the systematic \deleted{errors}\added{uncertainties} 
associated with binarity are
on the order of a few percent.

@@

For planets smaller than 2\,$R_\oplus$, the true occurrence rate
density is more uncertain because such planets are more difficult to
detect.  To investigate the systematic \deleted{errors}\added{uncertainties} 
associated with binarity
on the inferred occurrence rate of Earth-sized planets, we need to
make plausible assumptions about the true occurrence rate density.
Obviously if the rate density continues to vary as $r^{-3}$ down to
much smaller planet sizes, the results of the preceding section will
hold.  But there is no particular reason to think this will be the
case and indeed, some investigators have concluded that the rate
density begins to level off to a constant value as $r$ decreases below
$2R_\oplus$~\citep{petigura_plateau_2013}.  To investigate the
implications we consider a broken power-law:
\begin{align}
    f(r)
    &\propto
    \left.
    \begin{cases}
        r^\gamma & \text{for } r\geq 2R_\oplus \\
        {\rm constant} & \text{for } r\leq2R_\oplus.
    \end{cases}
    \right.
    \label{eq:model3_radius_distribution}
\end{align}
In this case the integrals that appear in
Equation~\ref{eq:general_Gamma_a} are tedious to work out
analytically, leading us to evaluate them numerically.\footnote{We
refer the interested reader to our online code for performing this
computation: \url{github.com/lgbouma/binary_biases}, commit
\texttt{fd41b4b}.} Figure~\ref{fig:occ_rate_model_3_log} shows the
results, again for the case $n\b/n\s=0.79$, $\alpha = 3.5$, $\beta=0$,
and $\gamma=-2.92$.

@@

The occurrence rate calculations that led to this discovery \deleted{did not
carefully account for}\added{gave only a preliminary accounting for} 
unresolved binaries.
Intuitively we expect
unresolved binaries to have a blurring effect, filling in any gaps in
the \added{true} radius distribution and making them appear less empty than in
reality.  To investigate the quantitative effects we make identical
assumptions as in Section~\ref{sec:model_3}, except that \deleted{the radius
distribution is assumed to have}\added{in analogy with
\citet{fulton_california-_2017}'s findings we assume the true radius 
distribution has} a complete absence of planets with
sizes between 1.5 and 2\,$R_\oplus$:
\begin{align}
    f(r)
    &\propto
    \left.
    \begin{cases}
        r^\gamma & \text{for } r\geq 2\,R_\oplus, \\
        0 & \text{for } 1.5\,R_\oplus < r < 2\,R_\oplus, \\
        {\rm constant} & \text{for } r\leq 1.5\,R_\oplus.
    \end{cases}
    \right.
    \label{eq:model4_radius_distribution}
\end{align}
Figure~\ref{fig:model_4} shows the resulting true and apparent
occurrence rate densities.  For the case $Z_2/Z_0=1$, unresolved
binaries reduce the contrast of the gap by nearly a factor of two,
while also producing spurious features for apparent sizes beneath
1.5\,$R_\oplus$.  Thus the gap identified by
\citet{fulton_california-_2017} may be even more devoid of planets
than it appears.

@@

Ignoring binarity introduces systematic errors to star and planet
counts in transit surveys, which in turn lead to biases in the derived
planet occurrence rates.  Thus far, occurrence rate calculations using
transit survey data have mostly ignored stellar
multiplicity~\citep[\textit{e.g.},][]{howard_planet_2012,fressin_false_2013,foreman-mackey_exoplanet_2014,dressing_occurrence_2015,burke_terrestrial_2015}.
We do not claim to have solved this problem. Our aim was to clarify
the various effects and assess the size of the 
\deleted{errors}\added{uncertainties} in various
situations. We hope that our formalism will also be useful to those
investigators who do attempt more realistic calculations of occurrence
rates.

@@

\paragraph{Earth-like planets} For the specific problem of calculating
the occurrence rate $\eta_\oplus$ of Earth-like planets with periods
of order one year, even a 50\% systematic error would not be the
dominant source of uncertainty at the moment.  Estimates by
\citet{youdin_exoplanet_2011},~\citet{petigura_prevalence_2013},~\citet{dong_fast_2013},~\citet{foreman-mackey_exoplanet_2014},
and~\citet{burke_terrestrial_2015} have found values of $\eta_\oplus$
ranging from $0.03$ to unity, with the wide variance arising from
small-number statistics as well as a host of other systematic
\deleted{errors}\added{uncertainties}~\citep[see][Figure~17]{burke_terrestrial_2015}.
  Rapid progress
is expected soon, though, and binarity will merit closer attention in
the future.  It may also be worth investigating the possible effects
of any dependence of occurrence rate densities on the period of the
binary.  Binaries with separations $\lesssim 10\,{\rm AU}$ could
provoke dynamical instabilities, leading to fewer Earth-like planets
per
star~\citep[\textit{e.g.},][]{holman_long-term_1999,wang_influence_2014,
kraus_impact_2016}.  This would affect transit survey measurements of
$\eta_\oplus$ beyond our rough estimate.

@@

\added{\paragraph{Caveats \& outlook}
Unresolved binaries corrupt occurrence rate distributions along
dimensions other than planet radius. 
In particular, since binarity can lead to incorrect
assessment of a planet's host stellar mass, the occurrence
distributions in semi-major axis and insolation must also be corrected.
We leave consideration of such biases for future work.}

\added{
It seems likely that to consistently account for binarity, one
must use a population inference model with parameters
controlling both the planet and binary star populations.
Given the difficulty of writing a correct likelihood function
for such a model, one feasible computational approach might be to
forward-model transit surveys with likelihood-free 
techniques~\citep[\textit{e.g.},][]{morehead_understanding_2016,hsu_improving_2018}.
The NASA {\it TESS} mission will likely open up this avenue of research,
since its target stars are close and bright.
}

>>>>

\title{ Biases in Planet Occurrence Caused by Unresolved Binaries in
        Transit Surveys }
%
\correspondingauthor{L. Bouma}
\email{luke@astro.princeton.edu}
%
\author[0000-0002-0514-5538]{L. G. Bouma}
\affiliation{
    Department of Astrophysical Sciences,
    Princeton University,
    4 Ivy Lane, Princeton, NJ 08540, USA}
\author[0000-0003-1298-9699]{K. Masuda}
\affiliation{
    Department of Astrophysical Sciences,
    Princeton University,
    4 Ivy Lane, Princeton, NJ 08540, USA}
\affiliation{
    NASA Sagan Fellow}
\author[0000-0002-4265-047X]{J. N. Winn}
\affiliation{
    Department of Astrophysical Sciences,
    Princeton University,
    4 Ivy Lane, Princeton, NJ 08540, USA}
%
%
\begin{abstract}
%
Wide-field surveys for transiting planets, such as the NASA {\it
Kepler} and {\it TESS} missions, are usually conducted without knowing
which stars have binary companions. Unresolved and unrecognized
binaries give rise to systematic errors in planet occurrence rates,
including misclassified planets and mistakes in completeness
corrections.  The individual errors can have different signs, making
it difficult to anticipate the net effect on inferred occurrence
rates. Here we use simplified models of signal-to-noise limited
transit surveys to try and clarify the situation. We derive a formula
for the apparent occurrence rate density measured by an observer who
falsely assumes all stars are single. The formula depends on the
binary fraction; the mass function of the secondary stars; and the
true occurrence of planets around primaries, secondaries, and single
stars. It also takes into account the Malmquist bias by which binaries
are over-represented in flux-limited samples. Application of the
formula to an idealized {\it Kepler}-like survey shows that for
planets larger than 2\,$R_\oplus$, the net systematic error is of
order 5\%.  In particular, unrecognized binaries are unlikely to be
the reason for the apparent discrepancies between hot Jupiter
occurrence rates measured in different surveys.  For smaller planets
the errors are potentially larger: the occurrence of Earth-sized
planets could be overestimated by as much as 50\%. We also show that
whenever high-resolution imaging reveals a transit host star to be a
binary, the planet is usually more likely to orbit the primary star
than the secondary star.
%
\end{abstract}
%
\keywords{
    methods: data analysis ---
    planets and satellites: detection ---
    surveys}
%
%

@@

A lingering concern about these studies is that in most cases,
investigators have assumed that all of the targets in the survey are
single stars~\citep[\textit{e.g.},][]{
howard_planet_2012,fressin_false_2013,
dressing_occurrence_2015,burke_terrestrial_2015}.  In reality, many of
the sources that are monitored in a transit survey are unresolved
multiple-star systems, mainly binaries.  Unrecognized binaries cause
numerous systematic errors in the planetary occurrence rates.  For
example, when there is a transiting planet around a star in a binary,
the additive constant light from the second star reduces the
fractional loss of light due to the planet.  This makes transit
signals harder to detect and lowers the number of detections.  On the
other hand, a binary system presents two opportunities to detect
transiting planets, which could increase the overall number of
detections.

@@

There are many potential pitfalls in this calculation.  Some genuine
transit signals are missed even if they formally exceed the
signal-to-noise threshold, because of the probabilistic nature of
transit detection.  Planets can be misclassified due to statistical
and systematic errors in the catalogued properties of the stars.  Some
transit-like signals are spurious, arising from noise fluctuations or
failures of ``detrending'' the astrophysical or instrumental
variations in the photometric signal.  Poor angular resolution leads
to blends between eclipsing binary stars and other stars along nearly
the same line of sight, producing signals that mimic those of
transiting planets.

Here, though, we will focus exclusively on problems that arise from
the fact that many stars exist in unresolved binary systems.  We will
also focus on the errors in planet occurrence as a function of radius,
rather than orbital period. This is because when more than one transit
is detected (as is usually required by the surveyors), the orbital
periods can be measured without ambiguity regardless of whether the
host star is single or one member of a binary.  Even with this narrow
focus, there are numerous sources of error.  All three of the
quantities in Equation~\ref{eq:occ_rate_simple} are biased:
\begin{enumerate}
%    
    \item The number of detected planets, $N_{\det}$, is actually the
      number of detected planets that {\it appear} to have size $r$,
      orbiting stars that {\it appear} to have mass $M$ \added{and
      radius $R$}.  Whenever the planet-hosting star is part of a
      binary,
%    
    \begin{itemize}
        \item the planet's size could be misclassified because of the
          reduction in the amplitude of the photometric signal;
%        
        \item the host star's properties could be misclassified
          because its light is combined with a second star of a
          different spectral type.
%        
    \end{itemize}
%    
    \item The number of searchable stars, $N_\star$, is biased
%    
    \begin{itemize}
%        
        \item toward lower values, because it does not include all of
          the secondary stars that were inadvertently searched for
          transiting planets;
%        
        \item toward higher values, because some of the stars that
          appeared to be searchable are in fact binaries for which the
          amplitude of the photometric signal would have been reduced
          to an undetectable level.
%        
    \end{itemize}
%    
    \item The transit probability $p_{\rm tra}$ is biased because the
      planet-hosting star could be misclassified.  At fixed orbital
      period, the transit probability is proportional to
      $\rho^{-1/3}$, where $\rho$ is the stellar mean
      density~\citep{Winn2010}.  Therefore, any errors in determining
      the host star's mean density lead to errors in the correction
      for the transit probability.
%    
\end{enumerate}

@@

We can now gauge the size of the systematic errors associated with
unresolved binaries, for \deleted{cases of interest.}\added{the cases
of interest listed in Table~\ref{tab:case_study_toc}.} Throughout this
section we assume $R\propto M$, and $L \propto M^{\alpha}$ where
$\alpha = 3.5$, as is roughly the case for main-sequence stars.  With
these assumptions, Equations~\ref{eq:dil_1} and~\ref{eq:dil_2} become
\begin{equation}
  \mathcal{D}_1
  =  (1+q^\alpha)^{1/2}~~{\rm and}~~\mathcal{D}_2
  = q(1+q^{-\alpha})^{1/2}.
\end{equation}

@@

Again assuming $\Gamma_0=\Gamma_1=\Gamma_2$, the apparent occurrence
rate density is 
\begin{align}
    %see annoying_q_factor_dec22.pdf for derivation
    \notag
    %\frac{\left.\Gamma\a(r\a)\right|_{r\a= r}}{\Gamma_0(r)} 
    \frac{\Gamma\a(r\a)}{\Gamma_0(r\a)} 
    &=
    \frac{1}{1+\mu}
    \left[
      1 + \frac{1}{\mathcal{N}_q} \frac{n\b}{n\s}
    \left(
      \int_0^1 {\rm d}q\,q^\beta (1+q^\alpha)^{\frac{\gamma+4}{2}} +
      \right.
      \right. \\
      %isn't latex great?
      &\quad\quad\quad\quad\quad\quad\quad\quad\quad\quad\quad
      \left.\left.
      \int_0^1 {\rm d}q\,q^{\beta+\gamma+\frac{5}{3}} 
      (1+q^\alpha)^{\frac{3}{2}}
      (1+q^{-\alpha})^{\frac{\gamma+1}{2}}
    \right)
    \right],
    \label{eq:powerlaw_vary_binary}
\end{align}
Adopting the realistic numerical values $n\b/n\s=0.79$, $\alpha =
3.5$, $\beta=0$, and $\gamma=-2.92$, the summed integrals in
Equation~\ref{eq:powerlaw_vary_binary} give $(\ldots)\approx 1.503$,
and $\Gamma\a/\Gamma_0 = 1.048$.  The apparent occurrence rate density
is 4.8\% higher than the true value.
\added{Exploring different values for the power-law index,
we find that $\Gamma\a/\Gamma_0$ lies between $0.99$ and 
$1.10$, for $\gamma=-3.5$ and $\gamma=-2.5$, respectively.}
We conclude that for planets around Sun-like stars with orbits within
0.25\,AU and radii in the range 2--17\,$R_\oplus$, the systematic
errors associated with binarity are on the order of a few percent.

@@

For planets smaller than 2\,$R_\oplus$, the true occurrence rate
density is more uncertain because such planets are more difficult to
detect.  To investigate the systematic errors associated with binarity
on the inferred occurrence rate of Earth-sized planets, we need to
make plausible assumptions about the true occurrence rate density.
Obviously if the rate density continues to vary as $r^{-3}$ down to
much smaller planet sizes, the results of the preceding section will
hold.  But there is no particular reason to think this will be the
case and indeed, some investigators have concluded that the rate
density begins to level off to a constant value as $r$ decreases below
$2R_\oplus$~\citep{petigura_plateau_2013}.  To investigate the
implications we consider a broken power-law:
\begin{align}
    f(r)
    &\propto
    \left.
    \begin{cases}
        r^\gamma & \text{for } r\geq 2R_\oplus \\
        {\rm constant} & \text{for } r\leq2R_\oplus.
    \end{cases}
    \right.
    \label{eq:model3_radius_distribution}
\end{align}
In this case the integrals that appear in
Equation~\ref{eq:general_Gamma_a} are tedious to work out
analytically, leading us to evaluate them numerically.\footnote{We
refer the interested reader to our online code for performing this
computation: \url{github.com/lgbouma/binary_biases}, commit
\texttt{fd41b4b}.} Figure~\ref{fig:occ_rate_model_3_log} shows the
results, again for the case $n\b/n\s=0.79$, $\alpha = 3.5$, $\beta=0$,
and $\gamma=-2.92$.

@@

The occurrence rate calculations that led to this discovery \deleted{did not
carefully account for}\added{gave only a preliminary accounting for} 
unresolved binaries.
Intuitively we expect
unresolved binaries to have a blurring effect, filling in any gaps in
the \added{true} radius distribution and making them appear less empty than in
reality.  To investigate the quantitative effects we make identical
assumptions as in Section~\ref{sec:model_3}, except that \deleted{the radius
distribution is assumed to have}\added{in analogy with
\citet{fulton_california-_2017}'s findings we assume the true radius 
distribution has} a complete absence of planets with
sizes between 1.5 and 2\,$R_\oplus$:
\begin{align}
    f(r)
    &\propto
    \left.
    \begin{cases}
        r^\gamma & \text{for } r\geq 2\,R_\oplus, \\
        0 & \text{for } 1.5\,R_\oplus < r < 2\,R_\oplus, \\
        {\rm constant} & \text{for } r\leq 1.5\,R_\oplus.
    \end{cases}
    \right.
    \label{eq:model4_radius_distribution}
\end{align}
Figure~\ref{fig:model_4} shows the resulting true and apparent
occurrence rate densities.  For the case $Z_2/Z_0=1$, unresolved
binaries reduce the contrast of the gap by nearly a factor of two,
while also producing spurious features for apparent sizes beneath
1.5\,$R_\oplus$.

\added{The quantitative results of this numerical experiment obviously
depend on the input assumptions.  The more general point is that
whenever there is a gap-like feature in the true rate density
distribution, the effects of binarity will tend to fill it in.  If a
gap is observed in an apparent distribution, the true distribution
must have a gap at least as deep, if not deeper.}  Thus, the gap
identified by \citet{fulton_california-_2017} may be even more devoid
of planets than it appears.

@@

Ignoring binarity introduces systematic errors to star and planet
counts in transit surveys, which in turn lead to biases in the derived
planet occurrence rates.  Thus far, occurrence rate calculations using
transit survey data have mostly ignored stellar
multiplicity~\citep[\textit{e.g.},][]{howard_planet_2012,fressin_false_2013,foreman-mackey_exoplanet_2014,dressing_occurrence_2015,burke_terrestrial_2015}.
We do not claim to have solved this problem. Our aim was to clarify
the various effects and assess the size of the errors in various
situations. We hope that our formalism will also be useful to those
investigators who do attempt more realistic calculations of occurrence
rates.

@@

\paragraph{Earth-like planets} For the specific problem of calculating
the occurrence rate $\eta_\oplus$ of Earth-like planets with periods
of order one year, even a 50\% systematic error would not be the
dominant source of uncertainty at the moment.  Estimates by
\citet{youdin_exoplanet_2011},~\citet{petigura_prevalence_2013},~\citet{dong_fast_2013},~\citet{foreman-mackey_exoplanet_2014},
and~\citet{burke_terrestrial_2015} have found values of $\eta_\oplus$
ranging from $0.03$ to unity, with the wide variance arising from
small-number statistics as well as a host of other systematic
errors~\citep[see][Figure~17]{burke_terrestrial_2015}.  Rapid progress
is expected soon, though, and binarity will merit closer attention in
the future.  It may also be worth investigating the possible effects
of any dependence of occurrence rate densities on the period of the
binary.  Binaries with separations $\lesssim 10\,{\rm AU}$ could
provoke dynamical instabilities, leading to fewer Earth-like planets
per
star~\citep[\textit{e.g.},][]{holman_long-term_1999,wang_influence_2014,
kraus_impact_2016}.  This would affect transit survey measurements of
$\eta_\oplus$ beyond our rough estimate.

@@

\added{\paragraph{Caveats \& outlook} In this paper we have focused
  exclusively on systematic errors in the planet radius distribution.
  Unresolved binaries will also have other effects on the inferred
  properties of a distribution of transiting planets.  For example,
  the combined light of a binary might lead to a mistaken estimate of
  the mass of the host star.  This would have implications for the
  occurrence rate as a function of orbital distance and insolation,
  since the orbital distance is usually calculated from the orbital
  period and the stellar mass.}

\added{Of course, many problems associated with binarity would
disappear if we had perfect knowledge about the companions to all
the stars in a transit survey. Progress toward this ideal is likely
to come soon, thanks to the combination of data from the ESA {\it
Gaia} astrometric mission \citep{Gaia2016} and the NASA {\it TESS}
mission~\citep{ricker_transiting_2014}.  The {\it TESS} transit
survey will concentrate on stars that are brighter than {\it Kepler}
stars by a factor of 30--100, and closer to the Earth by a factor of
5--10.  This will simplify all of the follow-up observations to
search for companions and measure their properties, with direct
imaging, Doppler monitoring, and astrometric monitoring.  Meanwhile
{\it Gaia} will deliver parallaxes, astrometic monitoring data, and
space-based apparent magnitudes for all of the stars relevant for {\it
TESS}, all of which will make it easier to identify binaries and other
multiple star systems.}
